beastscriptr: BEAUti for R

### 2017-10-26 article.tex
EOD

@@

  \caption{
    Workflow
  }
  \label{fig:beauti}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As in BEAUti, to create a BEAST2 input file, 
one needs an alignment, model priors and MCMC parameters.
When the alignment is stored in the file alignment.fasta,
the desired tree prior is a constant-rate Birth-Death 
model \cite{nee1994reconstructed} and the MCMC chain needs to run 1M 
states, beastscriptr can create a file 
(called \verb;birth_death.xml;) from R 
as such:

\begin{lstlisting}[language=R]
beast_scriptr(
  input_fasta_filename = "alignment.fasta",
  mcmc_chainlength = 10000000,
  tree_prior = "birth_death",
  output_xml_filename = "birth_death.xml"
)
\end{lstlisting}

The created file \verb;birth_death.xml; can then be used by BEAST2.

@@

\begin{lstlisting}[language=R]
beast_scriptr(
  input_fasta_filename = "alignment.fasta",
  mcmc_chainlength = 10000000,
  tree_prior = "birth_death",
  output_xml_filename = "birth_death_fixed_15.xml"
  fixed_crown_age = TRUE,
  initial_phylogeny = beastscriptr::fasta_to_phylo(
    input_fasta_filename, crown_age = 15)
)
\end{lstlisting}

>>>>

% Style of code
% From http://r.789695.n4.nabble.com/How-to-nicely-display-R-code-with-the-LaTeX-package-listings-tp4648110.html
\usepackage{fancyvrb} 
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\lstdefinestyle{mystyle}{
  language=R,% set programming language
  basicstyle=\ttfamily\small,% basic font style
  commentstyle=\color{gray},% comment style
  numbers=left,% display line numbers on the left side
  numberstyle=\scriptsize,% use small line numbers
  numbersep=10pt,% space between line numbers and code
  tabsize=2,% sizes of tabs
  showstringspaces=false,% do not replace spaces in strings by a certain character
  captionpos=b,% positioning of the caption below
  breaklines=true,% automatic line breaking
  escapeinside={(*}{*)},% escaping to LaTeX
  fancyvrb=true,% verbatim code is typset by listings
  extendedchars=false,% prohibit extended chars (chars of codes 128--255)
  literate={"}{{\texttt{"}}}1{<-}{{$\bm\leftarrow$}}1{<<-}{{$\bm\twoheadleftarrow$}}1
  {~}{{$\bm\sim$}}1{<=}{{$\bm\le$}}1{>=}{{$\bm\ge$}}1{!=}{{$\bm\neq$}}1{^}{{$^{\bm\wedge}$}}1,% item to replace, text, length of chars
  alsoletter={.<-},% becomes a letter
  alsoother={$},% becomes other
  otherkeywords={!=, ~, $, \&, \%/\%, \%*\%, \%\%, <-, <<-, /},% other keywords
  deletekeywords={c}% remove keywords 
}
\lstset{style=mystyle}

@@

  \caption{
    Workflow. From an alignment and priors, one creates a BEAST2 XML input file. This
    can be done using beastscript and BEAUti. The XML file created is run by BEAST2
    to create a posterior.
  }
  \label{fig:workflow}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As in BEAUti, to create a BEAST2 input file, 
one needs an alignment and model settings.
With those settings at their default, \verb;beastscriptr; is easy to use:
Create a BEAST2 input file with name \verb;beast2.xml; from 
a FASTA file with name \verb;alignment.fas; is done like this:

\begin{lstlisting}[language=R, caption=Simplest example]
library(beastscriptr)
create_beast2_input_file(
  "alignment.fas",
  "beast2.xml"
)
\end{lstlisting}

The created file \verb;beast2.xml; can then be used by BEAST2.

@@

\begin{lstlisting}[language=R, caption=Example with fixed crown age]
create_beast2_input_file(
  "alignment.fas",
  "beast2.xml"
  fixed_crown_age = TRUE,
  initial_phylogeny = beastscriptr::fasta_to_phylo(
    fasta_filename = "alignment.fas", 
    crown_age = 15
  )
)
\end{lstlisting}
